
```{r load, echo=FALSE}
library(pander)
library(stringr)
library(knitr)
library(jsonlite)
library(ggplot2)
library(readr)
source("r/utilities.R")
source("r/functions.R")
headverbs <- as.data.frame(fromJSON("../data/headverbs.json"))
genstats <- as.data.frame(fromJSON("../data/genstats.json"))
fsstats1 <- as.data.frame(fromJSON("../data/firstsentence.json"))

panderOptions('table.alignment.default', function(df) ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table',400)
#..pandoc("analysis.md",config="r/conf.txt")

#...Note: Taking only the first indicator into account:

fsstats1$asuminen_expressed <- gsub(";.*","",fsstats1$asuminen_expressed)
fsstats1$indicatorword <- gsub(";.*","",fsstats1$indicatorword)
fsstats1$head_of_indicator <- gsub(";.*","",fsstats1$head_of_indicator)
fsstats1$head_of_indicator_loc <- gsub(";.*","",fsstats1$head_of_indicator_loc)
fsstats1$indicatorloc <- gsub(";.*","",fsstats1$indicatorloc)


fsstats1$indicator_ratio  <- round(fsstats1$indicatorwords / fsstats1$words_total * 100,2)

fsstats1$headverb_person_simple <- "--"
fsstats1$headverb_person_simple[grepl("1",fsstats1$headverb_person)] <- "1p"
fsstats1$headverb_person_simple[grepl("3",fsstats1$headverb_person)] <- "3p"
fsstats1$headverb_person_simple[which(fsstats1$headverb_person=="")] <- "--"
fsstats1$indicatorword_length <-  str_length(fsstats1$indicatorword_token) 

write.csv(fsstats1,"updated_data.csv")

#fsstats1$paragraph[fsstats1$indicator_ratio==13.11]
#fsstats1$paragraph[fsstats1$indicator_ratio==2.90]

```

About the data
==============

* Paragraphs discussing *accommodation*

This theme was one of the topics suggested in the writing instructions for
the report. The part of the report that concerns accommodation is as follows:

    Asuminen: Oliko asunto valmiina saavuttuasi paikalle? Vuokrataso? Onko sinulla
    suosituksia asunnon hankinnassa tai muita neuvoja asumisen suhteen?


General statistical information about the data
==================================

* Total amount of texts: `r genstats["tekstit"]`
* Texts where accommodation as a topic not discussed: `r genstats["kpl_0"]`
* Texts with 1 paragraphs about accommodation: `r genstats["kpl_1"]`
* Texts with 2 paragraphs about accommodation: `r genstats["kpl_2"]`
* Texts with 3 paragraphs about accommodation: `r genstats["kpl_3"]`
* Texts with more than 3 paragraphs about accommodation: `r genstats["kpl_n"]`
* Total number of paragraphs about accommodation: `r genstats["asumiskappaleet"]`


Introducing the topic
=====================

First, let us take a look at how the topic (accommodation) is introduced to the
reader. 

Contrary to what at least used to be the case in English [cf. e.g. @dangelo86],
Finnish writing guides do not usually advice the writer to use special *topic
sentences*: sentences that summarize the contents of a paragraph and are most often
located at the beginning of it.

BTW, from [@smith2008braddock]:

    The third of Bain’s six rules called for a statement of topic in the initial
    sentence of the paragraph. Later scholars noted that the topic sentence could
    appear in another part of the paragraph or even be implied (Rogers, 1965).
    Furthermore, in his notable study, Braddock (1974) found that only 13% of the
    expository paragraphs in his study began with a topic sentence. Braddock’s
    study has been widely cited to support the view that reading and writing
    instruction needs to be modified to account for this fact.


However, even when there is no topic sentence *per se*, the reader must at some
point begin to have an idea about what the paragraph he is reading is about. 
When talking about accommodation, the most obvious way to indicate the topic
is, arguably, to use a word derived from the verb *asua*. Thus, we used the
asua-derivatives as *topic indicators* -- clues that indicate the reader
that the topic of the paragraph is accommodation. The data included
the following asua-derivatives:


```{r accommodationwords, echo=FALSE, results="asis"}

cat(paste("    ",paste(unique(fsstats1$indicatorword),sep=","),collapse=", "),"\n")

```


Interestingly enough, our data shows that out of the total `r nrow(fsstats1)`
texts that discuss accommodation `r table(fsstats1$sentence_number)["1"]`
(`r fn(table(fsstats1$sentence_number)["1"]/nrow(fsstats1)*100)` %)
use one of the aforementioned words in the **first sentence of the first paragraph that discusses 
accommodation.**


Let us now look at what these occurrences reveal about the paragraphs studied and the
writing strategies used by the authors.

Introducing strategies indicated by dependency roles
---------------------------------------------------

One way to get an idea about what kind of a strategy a writer is using to 
introduce the topic of accommodation is to look at the *dependency role*
given to the indicator word by the parser. The distribution of different dependency roles
of the topic indicators is given in table `r tabRef('tt_deproles')`:[^distexpl]

```{r accommodationwordstats, echo=FALSE, results="asis"}

do.words <- sort(table(fsstats1[fsstats1$asuminen_expressed=="dobj","indicatorword"]),decreasing=T)

tab <- table(fsstats1$asuminen_expressed[fsstats1$asuminen_expressed!="None"])
tab <- tab[tab>4]
tab <- tab[order(tab,decreasing=T)]
dftab <- as.data.frame(tab)
colnames(dftab) <- c("Dependency role","Frequency")
kable(dftab)

cat (tabRef('tt_deproles','Dependency roles of topic indicators'), "\n")

withindicator <- subset(fsstats1,asuminen_expressed!="None")
withindicator$indicator.deprel <- factor(withindicator$asuminen_expressed,levels=unique(names(sort(table(withindicator$asuminen_expressed),d=T))))

ggplot(withindicator,aes(x=indicator.deprel)) + geom_bar() + coord_flip() + theme_minimal()

```

[^distexpl]: The data in the table covers  the cases where the indicator
word is used in the first paragraph (in any sentence)


### Object

According to table `r tabRef('tt_deproles')` the most common dependency role is
the direct object. If we take a closer look at the cases which use a topic
indicator as the object of the first sentence, we can see that the majority of
them (`r fn(prop.table(do.words)["asunto"]*100)` %) use the word *asunto* ('apartment').
Further, if we look at the head verbs of these cases, the most frequent ones are 
the following:

```{r asuntoheads, echo=FALSE, results="asis"}
asunto<-subset(fsstats1,indicatorword=="asunto"&asuminen_expressed=="dobj")

asunto.heads <- sort(table(asunto$head_of_indicator),decreasing=T)[1:6]
asunto_hankkia<-subset(asunto,head_of_indicator=="hankkia")
asunto_hankkiaetc<-subset(asunto,head_of_indicator %in% names(asunto.heads))
indicator.placement.table <- sort(table(asunto_hankkiaetc$indicatorloc),decreasing=T)

do.indicator.placement.table <- sort(table(fsstats1[fsstats1$asuminen_expressed=="dobj","indicatorloc"]),decreasing=T)
do.indicator.placement.prob <- prop.table(do.indicator.placement.table)

pander(asunto.heads)
cat(tabRef("tt_asuntoheads","Head verbs of the topic indicator 'asunto' as object"),"\n")
```

This statistical information gives us a clear pattern of one popular strategy for 
starting a paragraph about accommodation. This strategy is exemplified by sentences
@ee_yliopistonkautta and @ee_yksityisilta:

(@ee_yliopistonkautta) Asunnon hankin yliopiston kautta.
(@ee_yksityisilta) Hankin asunnon yksityisiltä markkinoilta.

Example @ee_yliopistonkautta presents a variant with the topic indicator as the
first word; example @ee_yksityisilta uses the canonical verb-initial order.
Of all the `r sum(do.indicator.placement.table)` sentences where the topic indicator is a direct object, 
`r do.indicator.placement.table["1"]` (`r fn(100*do.indicator.placement.prob["1"])` %) are cases where 
the topic indicator is the first word like in example @ee_yliopistonkautta. 

### Root

As demonstrated in table `r tabRef('tt_deproles')`, direct objects are by far the most common
dependency role of topic indicators. The second dependency role with a
frequency well above all others is *root* -- that is, cases, where the topic is indicated
by the main verb of the sentence.

```{r, echo=FALSE}

deprels <- list(nmgobj=subset(fsstats1,asuminen_expressed=="nmod:gobj"),
                root=subset(fsstats1,asuminen_expressed=="root"),
                nsubj=subset(fsstats1,asuminen_expressed=="nsubj"),
                nmposs=subset(fsstats1,asuminen_expressed=="nmod:poss"),
                nscop=subset(fsstats1,asuminen_expressed=="nsubj:cop"),
                xcomp=subset(fsstats1,asuminen_expressed=="xcomp"),
                nmod=subset(fsstats1,asuminen_expressed=="nmod")
                )

rootverbs <- as.data.frame(sort(table(deprels$root$indicatorword),decreasing=T), optional=T)
subjects <- as.data.frame(sort(table(deprels$nsubj$indicatorword),decreasing=T), optional=T)

```

If we take a closer look at these cases, we can see, first of all, that
besides the most obvious option (the verb *asua*) the verbs that were
used included *majoittua* (`r rootverbs[rootverbs$Var1=="majoittua",2]` occurences)
*majoittaa* (`r rootverbs[rootverbs$Var1=="majoittaa",2]` occurences)
*vuokrata* (`r rootverbs[rootverbs$Var1=="vuokrata",2]` occurences)
and *alivuokrata* (`r rootverbs[rootverbs$Var1=="Alivuokrasimme",2]` occurences).

Further, the data reveals that of the `r nrow(deprels$root[deprels$root$indicatorword=="asua",])`
cases with *asua*, `r nrow(subset(deprels$root,indicatorword=="asua"&indicatorloc==1))` start with the verb.
The vast majority of these cases resemble example @ee_ulkopuolella in that they represent
the structure *asua* + time span + place of accommodation (the order of the
temporal and spatial arguments varies somewhat):

(@ee_ulkopuolella) Asuin koko lukuvuoden kampuksen ulkopuolella Brookfield Hallissa.

If the sentence does not begin with the verb, it is often the case that the adverbial of space
is the first constituent; another alternative is that the sentence begins with the pronoun *itse*
as in the following example:

(@ee_omassahuoneessa) Itse asuin yliopiston kampuksen asuntolassa omassa huoneessa.

The functions of these sentences are quite simple to describe: "During my stay at X I chose this kind of accommodation. 
Here's how it turned out to be: (positive / negative experiences)".

### Nsubj

Let us now look at the total `r nrow(deprels$nsubj)` cases which used a nominative subject to introduce the
topic. The most common subject was, unsurprisingly, *asunto* (`r subjects[1,2]` occurences). Others included 
*asuminen* (`r subjects[2,2]` occurences), asuntola (`r subjects[3,2]` occurences) and some unique cases such as 
*vuokrasopimus*.

Here are some examples of *asuminen*:

(@ee_asuminentapahtui) Asuminen tapahtui kampuksella asuntoloissa kahden henkilön huoneissa.
(@ee_asuminenjarjestyy) Asuminen järjestyy varsin kivuttomasti ja suhteellisen edullisesti kampuksella sijaitsevasta asuntolasta.

The examples with *asunto* represent a somewhat more heterogenous group.
If we look at the head verbs of these cases, we get the following table:

```{r asuntoverbs, echo=FALSE, results="asis"}

kable(as.data.frame(sort(table(deprels$nsubj$head_of_indicator[deprels$nsubj$indicatorword=="asunto"]),decreasing=T), optional=T))
cat(tabRef("tt_asuntoheads","Head verbs of the topic indicator 'asunto' as subject"),"\n")

```

The *olla*-cases can be grouped into a) possessive sentences like examples @ee_oliasunto and @ee_irlantiin:

(@ee_oliasunto) Minulla oli valmiiksi asunto, kun saavuin Prahaan.
(@ee_irlantiin) Minulla ei ollut asuntoa lähtiessäni Irlantiin.

and b) sentences like examples @ee_asuntovalmiina and @ee_asunnotvaihtareille:

(@ee_asuntovalmiina) Asuntoni oli valmiina heti saavuttuani paikalle.
(@ee_asunnotvaihtareille) Asunnot olivat vaihtareille valmiina, mutta saimme muuttaa niihin sisään vasta juuri ennen lukukauden alkua ja jouduimme lähtemää n niistä heti seuraavana päivänä lukukauden päättymisestä."

Other cases include the following:

(@ee_toas) Asunto järjestyi paikallisen Toas:in, Studentwerkin kautta.

Nmod:

[44] "Olimme jo Suomessa päättäneet olevamme hieman liian mukavuudenhaluisia ja suoraan sanottuna vanhoja jenkkien opiskelija-asuntoloihin."

Attempts at grouping the various ways of indicating the topic
------------------------------------------------------------

[check out this online whiteboard](https://app.ziteboard.com/?code=d977eb10-376e-4c84-a17a-ad7fac399ef0)


DOBJ
=====


```{r, echo=FALSE, results="asis"}

x <- withindicator$sentence[withindicator$indicator.deprel == "dobj"]


panderOptions('table.alignment.default', 'left')
panderOptions('table.alignment.rownames', 'left')
panderOptions('table.split.table', Inf)
panderOptions('table.emphasize.rownames', F)
#emphasize.italics.cols(1)
#pandoc.table(matrix(x,ncol=4))

dobj <- subset(withindicator,indicator.deprel=="dobj")
dobj <- ClassifyVerbs(dobj)
dobj$headverb_person[grepl("olin ",dobj$sentence, ignore.case=T) & dobj$headverb_person=="--"] <- "Sing.1"
sort(table(dobj$verbtype))
xtabs(~verbtype + headverb_person, data=dobj)

    


```








Kappaleen rajan käsittely
=========================

- Tsekkaa edellinen kappale, ja miten siinä johdateltu tulevaan aiheenvaihdokseen?
- Parempi: tsekataan tarpeen tullen -- tsekataan tietyiltä ryhmiltä







Literature
==========



